---
title: "code"
output: html_document
date: "2024-04-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE}
library(tidyverse)
library(MatchIt)
library(randomForest)
library(caret)
library(tidyverse)
library(VIM)
library(dplyr)
library(nnet)
library(ranger)
library(progress)

```


## Option 2: Explore a provided dataset for pre-specified research goals

I decided to go with *option c*, which involves constructing a propensity score to aid in an epidemiologic analysis of an etiologic question and then
performing an analysis using the propensity score to address that etiologic question.

My research question is as follows: How does length of hospital stay and subsequent patient disposition vary by type of insurance coverage in surgical patients?

After considering the nature of my research question, I considered the following variables to be key to answering my research question,

*Primary Variables:*
* Length of Stay (length_of_stay): The primary outcome variable.
* Patient Disposition (patient_disposition): The secondary outcome variable indicating what happens to patients after surgery (e.g., discharged to home, transferred to another facility).
* Type of Insurance Coverage (payment_typology_1, payment_typology_2, payment_typology_3): The primary explanatory variable. This indicates the insurance coverage types of the patients.

*Control Variables:*
* Age Group (age_group): Older patients might stay longer in hospitals due to complications.
* APR Severity of Illness (apr_severity_of_illness_description): This might influence both the length of stay and patient disposition.
* APR Risk of Mortality (apr_risk_of_mortality): This is crucial as it likely correlates with both the severity of the illness and the outcomes.
* Type of Admission (type_of_admission): Emergency, urgent, elective, etc., could affect both the length of stay and the type of disposition.
* APR Medical Surgical Description (apr_medical_surgical_description): Whether the admission was for a surgical procedure or a medical condition could impact the length of stay and disposition.


### Loading and pre-cleaning data

I converted the relevant variables into factors and ordered them, if applicable.
```{r data_prep}
hospital_df = read_csv("./data/sparcs_hospital.csv") %>% janitor::clean_names() %>%  select("payment_typology_1", "age_group", "apr_severity_of_illness_description", "apr_risk_of_mortality", "type_of_admission", "apr_medical_surgical_description", "patient_disposition", "length_of_stay")


# Coding as factors
hospital_df$payment_typology_1 <- as.factor(hospital_df$payment_typology_1)
hospital_df$type_of_admission <- as.factor(hospital_df$type_of_admission)
hospital_df$apr_medical_surgical_description <- as.factor(hospital_df$apr_medical_surgical_description)

# Coding as ordered factors
hospital_df$apr_severity_of_illness_description <- factor(hospital_df$apr_severity_of_illness_description, 
                                                   levels = c("Minor", "Moderate", "Major", "Extreme"), 
                                                   ordered = TRUE)

hospital_df$apr_risk_of_mortality <- factor(hospital_df$apr_risk_of_mortality, 
                                     levels = c("Minor", "Moderate", "Major", "Extreme"), 
                                     ordered = TRUE)

hospital_df$age_group <- factor(hospital_df$age_group, 
                                     levels = c("0 to 17", "18 to 29", "30 to 49", "50 to 69", "70 or Older" ), 
                                     ordered = TRUE)
```

My next step was to check for missingness in the variables of interest. Since the missing values were <0.005% of the dataset, the missing values were imputed via the mode.
```{r imputation}
# Specify variables to count missing values for
imp_vars <- c("length_of_stay", "patient_disposition",
                   "age_group", "apr_severity_of_illness_description",
                   "apr_risk_of_mortality", "type_of_admission", 
                   "apr_medical_surgical_description")

# Count missing values for specified variables
missing_counts <- hospital_df %>%
  select(one_of(imp_vars)) %>%
  summarise_all(~ sum(is.na(.)))

# Print the counts of missing values
print(missing_counts)

hospital_df<-na.omit(hospital_df)   

```


### Step 3: Estimate the Propensity Score using Multinomial Logistic Regression 
```{r}

multinom_model <- multinom(payment_typology_1 ~ age_group + apr_severity_of_illness_description + 
                           apr_risk_of_mortality + type_of_admission + apr_medical_surgical_description, 
                           data = hospital_df)
prop_score_multinom <- predict(multinom_model, type = "probs")

```


### Step 4: Estimate the Propensity Score using Random Forest
```{r randomforest}
set.seed(123)

# Define the trainControl
train_control <- trainControl(
  method = "cv",
  number = 5,
  verboseIter = TRUE
)

formula_rf <- payment_typology_1 ~ age_group + apr_severity_of_illness_description + 
           apr_risk_of_mortality + type_of_admission + apr_medical_surgical_description

# Calculate the number of predictor variables
num_predictors <- 5

# Create a revised grid that respects the number of predictors
feat.count <- c(num_predictors, num_predictors/2, sqrt(num_predictors))
grid.rf <- expand.grid(mtry=feat.count)

# Print the tuning grid to confirm the values
print(grid.rf)

# Adjust the number of trees to something manageable initially
tree.num <- seq(100, 300, by=100)  

# Loop through number of trees to optimize model
results.trees <- list()
for (ntree in tree.num) {
  set.seed(123)
  rf.train <- train(formula_rf,
                    data=hospital_df,
                    method="ranger",
                    trControl = train_control,
                    tuneGrid = grid.rf,
                    importance = 'impurity',
                    num.trees = ntree)  # specify number of trees directly in the train function
  index <- as.character(ntree)
  results.trees[[index]] <- rf.train
}

# Combine results and extract the best model parameters
output.trees <- do.call(rbind, lapply(results.trees, function(x) x$results))
best.tune <- output.trees[which.max(output.trees$Accuracy),]


# Fitting the final model with optimal parameters using ranger directly
set.seed(123)
ps.model.rf <- ranger(formula_rf,
                      data=hospital_df,
                      mtry=best.tune$mtry,
                      splitrule=best.tune$splitrule,
                      min.node.size=best.tune$min.node.size,
                      num.trees=best.tune$ntrees,
                      importance = 'impurity',
                      probability = TRUE)  # to get predicted probabilities

# Append predicted probabilities onto the original dataset
# Assuming binary classification or adjust accordingly for multinomial outcomes
hospital_df$PS.RF <- predict(ps.model.rf, data = hospital_df, type="response")$predictions[,2]

# Combine propensity scores with the original data
hospital_df$propensity_score_multinom <- prop_score_multinom[, "Insurance ype"]

# Fit multivariate regression model
fit <- svyglm(cbind(length_of_stay, patient_disposition) ~ payment_typology_1 + payment_typology_2 + payment_typology_3 + age_group + apr_severity_of_illness_description + apr_risk_of_mortality + type_of_admission + apr_medical_surgical_description + propensity_score_multinom + ps_rf, design = svydesign(ids = ~1, data = hospital_df))
summary(fit)


```
